Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning

### 2013-10-09 01_intro.tex
intro/conclusion

The outline of the paper is as follows. We first introduce the simple referential communication game investigated in \cite{Frank} and the original RSA model proposed there. We then analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics. Next we report the result of an experiment similar to the original one, and compare the predictions of different models to the new data. Finally we report on an additional experiment that suggests the sparsity of higher-order inference in actual human pragmatic reasoning.

>>>>

The outline of the paper is as follows. We first introduce the simple referential communication game investigated in \cite{Frank} and the original RSA model proposed there. Next we analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics. Finally we report on the result of an experiment similar to the original one, and compare the predictions of different models to the new data.

### 2013-10-09 06_conclusion.tex
intro/conclusion

Our main contributions are as follows: (1) We emphasized the distinction between belief, goal and action, systematically examined the RSA model, connected it to the game-theoretic approaches. (2) We tested the RSA model with a family of alternatives on the forced choice experiment and drew attention to the speaker's preference. Our data seems to suggest a different story about common knowledge --- what is commonly known is the Gricean pragmatic components, while both sides have his own private bias unknown to the other. (3) We explored the possibility of speakers doing second-order pragmatic inference and showed that it seems hard for them to do so.

Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to further improvement in predictive power. Of course, more data from a wide range of further experiments are needed for a full understanding of pragmatic inference about referential expressions.

>>>>

Our experiment and model comparison suggest the following: (1) on the conceptual level, 
it is helpful to clarify the distinction between belief, goal and action, and by systematically examining each component in the formal model we can more explicitly spell out the underlying assumptions and formulate alternatives to be tested. (2) on the empirical level, we tested the RSA model with a family of alternatives on a forced-choice experiment and investigated the speaker's preference as well as the listener's perceptual salience. Our data seem to suggest a different story about common knowledge --- what is commonly known is the Gricean pragmatic components, while both sides have his own private bias unknown to the other.

Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to a better understanding of the phenomenon and further improvement in the formal model's predictive power. Of course, more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.
